---
title: "PA1_template.Rmd"
author: "Dylan van den Berg"
date: "Tuesday, July 14, 2015"
output: html_document
---
##Read in the Data
```{r readdata}
##Ensure to set working directory to directory with data
data<- read.csv("activity.csv")

##Create a time sequence based on the interval
time<-paste(data[,2],"-",floor(data[,3]/100),":",(data[,3] - floor(data[,3]/100)*100), sep="")
time<-strptime(time,"%Y-%m-%d-%H:%M" )
time_interval<- paste(floor(data[,3]/100),":",(data[,3] - floor(data[,3]/100)*100), sep="")
time_interval<-strptime(time_interval,"%H:%M" )
data<-cbind(data,time,time_interval)
```
##What is mean total number of steps taken per day?
Calculate the steps per day
```{r stepsperday}
Steps_per_day<- aggregate(steps ~ date, data = data, sum, na.rm = TRUE)
```
Plot a histogram of the steps per day
```{r stepsperdayplot}
 hist(as.numeric(Steps_per_day[,2]), main = "Steps per day", xlab = "Steps")
```
Calculate and report the mean and the median (SPD)
```{r stepsperdaymeanmedian}
mean(Steps_per_day[,2], na.rm = TRUE)
median(Steps_per_day[,2], na.rm = TRUE)
```

##What is the average daily activity pattern?
```{r stepsbyinterval}
Steps_per_interval<- aggregate(steps ~ time_interval, data = data, mean, na.rm = TRUE)
##Note that the time interval is used and not the original interval
plot(x = Steps_per_interval[,1], y = Steps_per_interval[,2], type = "l", xlab = "Time Interval", ylab = "Average Number of Steps per Time Interval", main = "Average Steps per Time Interval")
```
Maximum steps in a single time interval (average):
```{r maxaveragesteps}
max(Steps_per_interval[,2], na.rm = TRUE) ##What is the max. average steps?
which.max(Steps_per_interval[,2]) ##Which time interval is the max located?
```

##Imputing missing values
Total Number of missing values:
```{r missingvaluescount}
na_test<-is.na(data[,1])
sum(na_test) ##Number of missing values
```
Method for inputing data - use average for time interval.

```{r replaceNAwithintervalave}
##Step through each observation and check if NA
##If NA, check for value in Steps_per_interval
##create new_data set
new_data<- data
Steps_per_interval_notime<- aggregate(steps ~ interval, data = data, mean, na.rm = TRUE)
for (i in 1:length(new_data[,1])){   ##Stepping through each steps value
        if (na_test[i] == TRUE){   ##If a NV value is found then it is changed 
                if(length(which(Steps_per_interval_notime == new_data[i,3]))> 1){
                        ##Above code checks if steps value is repeated (only repeated for 0)
                    new_data[i,1]<- 0  ##There are more than one 0 value, this corrects for it
                    }
                 else{
                new_data[i,1]<- Steps_per_interval_notime[which(Steps_per_interval_notime == new_data[i,3]),2]
                }
}
}
```
Calculate the New steps per day
```{r Newstepsperday}
New_Steps_per_day<- aggregate(steps ~ date, data = new_data, sum, na.rm = TRUE)
```
Plot a histogram of the new steps per day
```{r newstepsperdayplot}
 hist(as.numeric(New_Steps_per_day[,2]), main = "New Steps per day", xlab = "Steps")
```
Calculate and report the mean and the median (SPD) of New data
```{r newstepsperdaymeanmedian}
mean(New_Steps_per_day[,2], na.rm = TRUE)
median(New_Steps_per_day[,2], na.rm = TRUE)
```
##Are there differences in activity patterns between weekdays and weekends?
```{r dateasfactorplot}
##Must install: install.packages("timeDate")
library(timeDate)
##Add a factor variable for weekdays or weekends
weekends<- isWeekday(new_data$date)
days<-weekdays(as.Date(new_data[,2]))
f_data<-cbind(new_data,days)
f_data<-cbind(f_data, weekends)
f_data$weekends<- factor(f_data$weekends, levels = c(FALSE,TRUE), labels = c('weekend','weekday'))

attach(f_data)
weekend_ave<- aggregate(steps, by=list(interval, weekends), data = f_data, FUN = mean)
names(weekend_ave)<-c("interval","weekends","ave_steps")
weekend_ave$weekends<-factor(weekend_ave$weekends, levels = c(FALSE, TRUE), labels = c('weekend','weekday'))
##Copying over the time series for each time interval:
 weekend_ave<-cbind(weekend_ave,f_data[1:576,5])
names(weekend_ave)<-c("interval","weekends","ave_steps","time_interval")
library(lattice)
xyplot(ave_steps ~ time_interval | weekends, data = weekend_ave, type = "l", layout = c(1,2), xlab = ("Time Interval"),main = "Time interval shown in hours from 00:00. Date of 17 July is arbitrary")

```


End of file
